---
title: "Exploratory Analysis of Copora for Text Mining"
author: "Ben Daniel"
date: "March 15, 2016"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(NLP)
library(tm)
library(reader)
library(wordcloud)
library(RTextTools)
library(RWeka)
library(slam)
library(SnowballC)
```
# Introduction
Text Mining and Predictive Text Analytics are interesting areas of study and application in the modern world of data science.  Their applications are numerous including forensics, marketing, and customer relationship management.  Even mobile devices leverage text mining as keyboards predict the word a user is about to type in order to improve text entry.

The Capstone Project for the Data Science Certificate offered by Johns Hopkins University requires students to write a prediction model for text using the principles learned in the certification courses. This report is an analysis of the corpora (the text samples) given for building the prediction algorithm.  It will suvery the documents, provide summary information about them and discuss intersting findings.  Additionally, it will cover the sampling and modeling strategy to complete the modeling assignment.

```{r echo=FALSE}
        #get file if it does not already exist and unzip contents
        fileName <- "https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip"
        destFile <- "data/Coursera-SwiftKey.zip"
        
        #set code for processing files into vectors (this may be an input into the shiny app at the end)
        lang_code <- "en_US"
        #data source type (twitter, etc.) -- this also could be an input from the shiny app
        dataSource<-".twitter.txt"  #other choice includes news.txt and blogs.txt
        #load profanity for filtering
        con_bw<-file("data/bad-words.txt")
        profanity<-readLines(con_bw)
        close(con_bw)
        
        #set the seed to enable reproducible research
        set.seed(1234)
        
        if(!file.exists("data/Coursera-SwiftKey.zip"))
        {
                download.file(fileName,destfile = destFile)
                unzip("data/Coursera-SwiftKey.zip", junkpaths = TRUE,overwrite = TRUE, exdir="data")
        }
        #make sure to get the bad-words file 
        if(!file.exists("data/bad-words.txt")){
                download.file("http://www.cs.cmu.edu/~biglou/resources/bad-words.txt",destfile="data/bad-words.txt")
        }
        #read in the profanity base
        con_bw<-file("data/bad-words.txt")
        profanity<-readLines(con_bw)
        close(con_bw)
        
        #load data
(copora <- VCorpus(DirSource("data/final/en_US/"),readerControl=list(language="en")))
```
#Data Survey
First, lets look at the data given to us.  Reading in the directory of text samples, we see the following:
```{r echo=TRUE}
metaCopora<-lapply(copora[1:length(copora)],meta)
```
